Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-10 appendix.tex
updated iso discussion in appendix

\section{Linear problem in isothermal limit}\label{iso_discuss}  
Here we examine in more detail the special case of isothermal
perturbations ($\beta\equiv 0$) in the VGSB framework.   
%Isothermal perturbations correspond to $\beta\equiv0$ ($\chi=1$) and 
In this case Eq. \ref{ode_Q} becomes   
\begin{align} 
  Q = W 
\end{align}
(i.e. $\delta P = c_s^2\delta \rho$), and so $  \ii\hat{\sigma}c_s\delta v_z = W^\prime $
from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact
simpler to work with an equation for $W$ by substituting $Q=W$ into
Eq. \ref{ode_w}. We obtain 
\begin{align}
  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
\end{align}
where for discussion purposes we have defined $f(\zhat)$ such that
\begin{align}\label{fz_shear}
  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
\end{align}
By comparing Eq. \ref{fz_shear} with Eq. \ref{vertical_shear_ex}, we
see that $f(\zhat)= 
\zhat\left(1+\epsilon^2\zhat^2\right)^{-3/2}$. More generally, though,
$f$ may be considered as a representation of the vertical
shear profile. Physically, we expect there is a maximum value of 
$|d\Omega/dz|$. The existence of a maximum, as remarked in
\S\ref{vshear_def}, limits the growth rate of the VSI. We explicitly
demonstrate this below.

\subsection{Necessity of vertical shear for
  instability}\label{integral_relation} 
We can establish a necessary condition for instability by multiplying
Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
terms when integrating by parts, by assuming $W$ or
$W^\prime$ vanishes at the boundaries, or that the boundaries are 
sufficiently far away so that the boundary terms are negligible because of the
decaying background density with increasing height. Then,
\begin{align}
  \hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat
  =\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
\end{align}
It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
shear. A similar conclusion can be reached in the equivalent linear  
problem in global cylindrical disk geometry.

\subsection{Maximum growth rate}   
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
% \ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, 
The real and imaginary parts of 
Eq. \ref{integral_relation1} are
\begin{align}
  \left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
  \notag
  =\real\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\zhat\right], \\
   2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}
  =\imag\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right],
\end{align}
where we recall $\hat{\omega} = \real \hat{\sigma}$ and
$\hat{\nu}=\imag\hat{\sigma}$. 
Adding the square of these equations give
\begin{align}
  \left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2
  +4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}
  =\left|\ii
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
  4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime
  \right|^2d\hat{z} 
  \leq\left|
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  \left( \int_{\zhat_1}^{\zhat_2}\rho
    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
  \int_{\zhat_1}^{\zhat_2}\rho 
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
  \left|f(\hat{z})W^*W^\prime \right|d\hat{z}
  \leq
  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
  |W|\left|W^\prime \right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  |\hat{\nu}|\leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
vertical shear in the domain considered.

@@

\subsection{Explicit solutions in the thin-disk limit}
For thin disks ($\epsilon\ll1$), we write 
$\ln\rho\simeq -\zhat^2/2$ and $f(\zhat)\simeq \zhat$, 
% by expanding the background density and vertical shear profile in powers
% of $z/r$. To lowest order we obtain 
% \begin{align}
%   &\frac{\rho}{\rho_0} \simeq 
%   \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
%   &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
% \end{align} 
%With this approximation and those in \S\ref{integral_relation}, 
so Eq. \ref{iso_ode1} becomes  
\begin{align}\label{iso_ode3}
  W^{\prime\prime} - \left(1 + \ii q\epsilon
    \hat{k}\right)\hat{z}W^\prime  +
  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
  0.
\end{align}
We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
that derived by \cite{nelson13}, although we have taken a different
route.  
% To complete the problem we must specify boundary conditions. The 
% maximum vertical domain size should be limited by the thin
% disk and low-frequency approximations (i.e. $z\lesssim r$).     
%It is, however, common to take $\mathrm{max}|z|\to\infty$ and 
We impose the kinetic energy remain bounded at infinity. Even though
we have adopted the thin-disk approximation, an infinite
disk is appropriate in the sense that both the true density profile 
and its thin-disk approximation decay rapidly away from the midplane.

However, the thin-disk representation of vertical shear diverges with 
$\hat{z}$, unlike the true vertical shear profile which eventually
decays. Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit 
unphysically large growth rates (as demonstrated below). Nevertheless,
the infinite disk is useful to consider since it permits an analytical
discussion.

\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
When $q=0$, Eq. \ref{iso_ode3} is
\begin{align}\label{hermite_ode}
  \left[w(\hat{z})W^\prime \right]^\prime + nW
  w(\hat{z}) =0, 
\end{align}
where $w(\zhat)=\exp{(-\zhat^2/2)}$ and we have defined a new eigenvalue 
\begin{align}
  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
\end{align} 
Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
that the kinetic energy density remain bounded at infinity, then the
solutions are Hermite polynomials  
\begin{align}
  W \propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to a real
eigenfrequency $\sigma$ such that
\begin{align}
  \left|\hat{\sigma}\right| = \sqrt{n}
  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
n-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height. Note that
Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
\cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
nearly-Keplerian approximation in that equation.

We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
has also been qualitatively observed in numerical simulations
\citep{stoll14}.

\subsubsection{Polynomial solutions with vertical
  shear}\label{iso_poly}
In the presence of vertical shear ($q\neq0$) we can still seek power-series solution to
Eq. \ref{iso_ode3}, 
\begin{align}
  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
\end{align}
Then the coefficients must satisfy the recurrence relation
\begin{align}
  (l+2)(l+1)a_{l+2} +
  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
\end{align}
where $n$ is not necessarily an integer. Indeed, if we demand
a polynomial of degree $L$ as the solution, then the eigenfrequency
must satisfy
\begin{align}\label{sig2_iso}
  \hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
      \hat{k}}{1+\hat{k}^2}\right).
\end{align}
The eigenfrequency $\hat{\sigma}$ is therefore complex for
$L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
derived by setting $\beta=0$ in the more general dispersion relation 
Eq. \ref{relax_disp} with coefficients given in Appendix
\ref{relax_coeff}. We are interested in 
unstable modes for which $\hat{\nu}>0$ and is given by
\begin{align}\label{simple_growth}
  \hat{\nu} =\sqrt{
    \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
      1\right)}. 
\end{align}
For fixed $q$ and $\epsilon$, the growth rate vanishes for both
$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
The maximum growth rate
occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
\begin{align}
  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
\end{align}
and
\begin{align}
  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
    q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
\end{align}
Since $L\geq1$, there is no limit to the maximum growth rate. This is
an artifact of the thin-disk approximation because the magnitude of vertical 
shear $|d\Omega^2/dz|\propto |z|$ increases indefinitely with 
height. However, Eq. \ref{iso_max_growth} compares well with
numerical solutions for small $L$ (see \S\ref{vertiso_pertiso}).

>>>>

\section{Linear problem in the isothermal limit}\label{iso_discuss}  
Here we examine in more detail the special case of isothermal
perturbations ($\beta\equiv 0$) for vertically isothermal disks 
($\Gamma=1$) in the VGSB framework.   
%Isothermal perturbations correspond to $\beta\equiv0$ ($\chi=1$) and 
In this case Eq. \ref{ode_Q} becomes   
\begin{align} 
  Q = W 
\end{align}
(i.e. $\delta P = c_s^2\delta \rho$), and so $  \ii\hat{\sigma}c_s\delta v_z = W^\prime $
from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact
simpler to work with an equation for $W$ by substituting $Q=W$ into
Eq. \ref{ode_w}. In this case, we shall not yet make the low-frequency
approximation, but first make the nearly-Keplerian approximation. We
obtain, in terms of dimensionless variables, 
\begin{align}
  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \frac{\ii \epsilon q \hat{k} f(\zhat)
      }{1-\hat{\sigma}^2} \right]W^\prime +
  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W,\label{iso_ode0} 
\end{align}
where for discussion purposes we have defined $f(\zhat)$ such that
\begin{align}\label{fz_shear}
  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
\end{align}
By comparing Eq. \ref{fz_shear} with Eq. \ref{vertical_shear_ex}, we
see that $f(\zhat)= 
\zhat\left(1+\epsilon^2\zhat^2\right)^{-3/2}$. More generally, though,
$f$ may be considered as a representation of the vertical
shear profile. Physically, we expect there is a maximum value of 
$|d\Omega/dz|$. The existence of a maximum, as remarked in
\S\ref{vshear_def}, limits the growth rate of the VSI. We explicitly
demonstrate this below.

\subsection{Stability conditions in the low-frequency limit}
In the low-frequency limit we assume $|\hat{\sigma}|\ll 1$ and
approximate Eq. \ref{iso_ode0} as 
\begin{align}
  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
    f(\zhat)\right]W^\prime +
  \hat{\sigma}^2\left(1+\hat{k}^2\right)W. \label{iso_ode1}
\end{align}
\subsubsection{Necessity of vertical shear for
  instability}\label{integral_relation} 
We can establish a necessary condition for instability by multiplying
Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
terms when integrating by parts, by assuming $W$ or
$W^\prime$ vanishes at the boundaries, or that the boundaries are 
sufficiently far away so that the boundary terms are negligible because of the
decaying background density with increasing height. Then,
\begin{align}
  \hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat
  =\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
\end{align}
It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
shear. A similar conclusion can be reached in the equivalent linear  
problem in global cylindrical disk geometry.

\subsubsection{Maximum growth rate}   
Here we show that the growth rate is limited by the maximum vertical
shear in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
% \ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, 
The real and imaginary parts of 
Eq. \ref{integral_relation1} are
\begin{align}
  \left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
  \notag
  =\real\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\zhat\right], \\
   2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}
  =\imag\left[\ii
    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right],
\end{align}
where we recall $\hat{\omega} = \real \hat{\sigma}$ and
$\hat{\nu}=\imag\hat{\sigma}$. 
Adding the square of these equations give
\begin{align}
  \left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2
  +4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}
  =\left|\ii
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
It is clear that
\begin{align}\label{sigma_finite_domain} 
  4\hat{\nu}^2\left(1+\hat{k}^2\right) 
  \int_{\zhat_1}^{\zhat_2}\rho
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime
  \right|^2d\hat{z} 
  \leq\left|
    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
\end{align}
On the left hand side of this inequality, we apply the Cauchy-Schwarz
inequality to obtain
\begin{align}
  \left( \int_{\zhat_1}^{\zhat_2}\rho
    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
  \int_{\zhat_1}^{\zhat_2}\rho 
  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
\end{align}
On the right hand side of Eq. \ref{sigma_finite_domain} we have
\begin{align}
  \left|\int_{\zhat_1}^{\zhat_2}\rho
    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
  \left|f(\hat{z})W^*W^\prime \right|d\hat{z}
  \leq
  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
  |W|\left|W^\prime \right|d\hat{z},
\end{align}
where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
Eq. \ref{sigma_finite_domain} gives
\begin{align}\label{max_growth}
  |\hat{\nu}|\leq
  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
\end{align}
Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
follows that the maximum possible growth rate of unstable modes,
satisfying the above boundary conditions, is limited by the maximum
vertical shear in the domain considered.

@@

\subsection{Explicit solutions in the thin-disk limit}\label{iso_explicit}
Here we assume a thin disk ($\epsilon\ll1$) so that $\ln\rho\simeq
-\zhat^2/2$ and $f(\zhat)\simeq \zhat$. However, we do not assume
low-frequency from the onset. Then Eq. \ref{iso_ode0} becomes 
\begin{align}\label{iso_ode3}
  W^{\prime\prime} - \left(1 + \frac{\ii q\epsilon \hat{k}}{1-\hat{\sigma}^2}\right)\hat{z}W^\prime  +
  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W = 0.
\end{align}
We remark that for $\hat{k}^2\gg 1$ and taking low-frequency limit,
Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in \cite{nelson13},
although we have taken a different route.

% To complete the problem we must specify boundary conditions. The 
% maximum vertical domain size should be limited by the thin
% disk and low-frequency approximations (i.e. $z\lesssim r$).     
%It is, however, common to take $\mathrm{max}|z|\to\infty$ and

We seek power series solutions to Eq. \ref{iso_ode3}, 
\begin{align}
  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
\end{align}
Then the coefficients must satisfy the recurrence relation
\begin{align}
  (l+2)(l+1)a_{l+2} +
  \left[\hat{\sigma}^2\left(1+\frac{\khat^2}{1-\hat{\sigma}^2}\right)
    - l\left(1+\frac{\ii \epsilon q
        \hat{k}}{1-\hat{\sigma}^2}\right)\right] a_l = 0. 
\end{align}
We demand the series to terminate  at $l=L$, i.e. a polynomial
solution, so that the vertical kinetic energy remain bounded as
$|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\sigma}$ is given via 
\begin{align}
\hat{\sigma}^4 - \left(L+1+\khat^2\right)\hat{\sigma}^2 + L\left(1 +
  \ii\epsilon q \khat\right) = 0. \label{simple_growth}
\end{align}

\subsubsection{Stability without vertical shear}
In the absence of vertical shear $q=0$, Eq. \ref{simple_growth} can be
written as \begin{align}
  \hat{\sigma}^2\khat^2 =
  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right). 
\end{align}
This is just the dispersion relation for axisymmetric waves in a
vertically isothermal disk already discussed in previous works 
\citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14}. In
this case the solutions are Hermite polynomials, $W\propto
\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma}$ is real so the disk
is stable, consistent with \S\ref{integral_relation}.

\subsubsection{Instability with vertical shear}
A simple solution to Eq. \ref{simple_growth} may be obtained by
considering small radial lengthscales $\khat\gg 1$. Then for the
low-frequency branch of Eq. \ref{simple_growth} 
\begin{align}
  \hat{\sigma}^2 \simeq L\left(\frac{1+\ii q \epsilon
       \hat{k}}{L+1+\hat{k}^2}\right).\label{simple_growth2}
\end{align}
The eigenfrequency $\hat{\sigma}$ is therefore complex for 
$L\geq1$, which signifies instability since we can choose
the sign of the square root to make $\hat{\nu} = \imag\hat{\sigma}>0$.

% \begin{align}\label{sig2_iso}
%   \hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
%       \hat{k}}{1+\hat{k}^2}\right).
% \end{align}
% The eigenfrequency $\hat{\sigma}$ is therefore complex for
% $L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
% derived by setting $\beta=0$ in the more general dispersion relation 
% Eq. \ref{relax_disp} with coefficients given in Appendix
% \ref{relax_coeff}. We are interested in 
% unstable modes for which $\hat{\nu}>0$ and is given by
% \begin{align}
%   \hat{\nu} =\sqrt{
%     \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
%       1\right)}. 
% \end{align}

% We impose the kinetic energy remain bounded at infinity. Even though
% we have adopted the thin-disk approximation, an infinite
% disk is appropriate in the sense that both the true density profile 
% and its thin-disk approximation decay rapidly away from the midplane.

% However, the thin-disk representation of vertical shear diverges with 
% $\hat{z}$, unlike the true vertical shear profile which eventually
% decays. Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit 
% unphysically large growth rates (as demonstrated below). Nevertheless,
% the infinite disk is useful to consider since it permits an analytical
% discussion.

% \subsubsection{Stability in the absence of vertical
%   shear}\label{iso_stable}
% \
% Eq. \ref{iso_ode3} is
% \begin{align}\label{hermite_ode}
%   \left[w(\hat{z})W^\prime \right]^\prime + nW
%   w(\hat{z}) =0, 
% \end{align}
% where $w(\zhat)=\exp{(-\zhat^2/2)}$ and we have defined a new eigenvalue 
% \begin{align}
%   n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
% \end{align} 
% Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
% that the kinetic energy density remain bounded at infinity, then the
% solutions are Hermite polynomials  
% \begin{align}
%   W \propto \He_n(\hat{z}),
% \end{align}
% and $n$ is a non-negative integer. This translates to a real
% eigenfrequency $\sigma$ such that
% \begin{align}
%   \left|\hat{\sigma}\right| = \sqrt{n}
%   \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
% \end{align}
% Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
% analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
% n-1$. Physically, this corresponds to radial length-scales much
% smaller than the local disk scale height. Note that
% Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
% \cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
% nearly-Keplerian approximation in that equation.

% We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
% Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
% has also been qualitatively observed in numerical simulations
% \citep{stoll14}.

% \subsubsection{Polynomial solutions with vertical
%   shear}\label{iso_poly}

% For fixed $q$ and $\epsilon$, the growth rate vanishes for both
% $\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
% The maximum growth rate
% occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
% \begin{align}
%   |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
% \end{align}
% and
% \begin{align}
%   \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
%     q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
% \end{align}
% Since $L\geq1$, there is no limit to the maximum growth rate. This is
% an artifact of the thin-disk approximation because the magnitude of vertical 
% shear $|d\Omega^2/dz|\propto |z|$ increases indefinitely with 
% height. However, Eq. \ref{iso_max_growth} compares well with
% numerical solutions for small $L$ (see \S\ref{vertiso_pertiso}).
